Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition

### 2018-12-03 paper/paper.Rmd
add chart and model effect models to paper

---
title: "Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
author: "Andreas Beger and Michael D. Ward"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts. The first project trial period took place in 2018, during the course of which volunteer participants, Amazon Mechanical Turkers, and an automated time series forecasting module we developed provided forecasts for more than 150 questions covering a diverse set of topics. We investigate two questions: (1) how does the performance of turker forecasters compare to volunteers, and (2) what impact did access to the machine forecasts have on forecaster accuracy? \\newline \\newline
"
bibliography: "../../whistle/master.bib"
header-includes:
  - 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

@@

|Forecaster | avg_Brier|     n|
|:----------|---------:|-----:|
|Machine    |     0.386|  1975|
|Turker     |     0.433| 39140|
|Volunteer  |     0.322|  7816|

Table: Average Brier score by forecaster group.

@@

Teasing out the chart and model effects is a bit harder through pairwise comparison.

Add in the LM and LMER model results here.

```
Coefficients:
                          Estimate Std. Error t value Pr(>|t|)    
(Intercept)               0.447126   0.002887 154.862  < 2e-16 ***
ForecasterVolunteer      -0.104321   0.006208 -16.805  < 2e-16 ***
IFP_GroupChart only       0.142377   0.017834   7.984 1.45e-15 ***
IFP_GroupChart and model -0.063318   0.008648  -7.322 2.48e-13 ***
sees_chartTRUE           -0.038067   0.015751  -2.417   0.0157 *  
sees_modelTRUE            0.042337   0.015251   2.776   0.0055 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 0.4776 on 46950 degrees of freedom
Multiple R-squared:  0.01339,	Adjusted R-squared:  0.01328 
F-statistic: 127.4 on 5 and 46950 DF,  p-value: < 2.2e-16
```

```
Linear mixed model fit by REML ['lmerMod']
Formula: brier ~ Forecaster + IFP_Group + sees_chart + sees_model + (1 |      ifp_id)
   Data: common_period[common_period$Forecaster != "Machine", ]

REML criterion at convergence: 50272.9

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-2.9464 -0.5432 -0.2071  0.3053  4.9013

Random effects:
 Groups   Name        Variance Std.Dev.
 ifp_id   (Intercept) 0.06386  0.2527  
 Residual             0.16808  0.4100  
Number of obs: 46956, groups:  ifp_id, 156

Fixed effects:
                          Estimate Std. Error t value
(Intercept)               0.438521   0.025392  17.270
ForecasterVolunteer      -0.077777   0.005519 -14.091
IFP_GroupChart only       0.160131   0.107529   1.489
IFP_GroupChart and model -0.061051   0.044711  -1.365
sees_chartTRUE           -0.021495   0.013689  -1.570
sees_modelTRUE            0.026931   0.013280   2.028

Correlation of Fixed Effects:
            (Intr) FrcstV IFP_Go IFP_am ss_cTRUE
FrcstrVlntr -0.042                              
IFP_GrpChro -0.237  0.029                       
IFP_GrpCham -0.567  0.005  0.140                
ss_chrtTRUE  0.011 -0.266 -0.099 -0.066         
ses_mdlTRUE -0.012  0.284  0.081 -0.016 -0.816  
```

@@

>>>>

---
title: "Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
author: "Andreas Beger and Michael D. Ward"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts. The first project trial period took place in 2018, during the course of which volunteer participants, Amazon Mechanical Turkers, and an automated time series forecasting module we developed provided forecasts for more than 150 questions covering a diverse set of topics. We investigate two questions: (1) how does the performance of turker forecasters compare to volunteers, and (2) what impact did access to the machine forecasts have on forecaster accuracy? \\newline \\newline
"
bibliography: "../../whistle/master.bib"
header-includes:
  - \usepackage{mathpazo}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library("readr")
library("dplyr")
library("knitr")
```

@@

```{r}
knitr::opts_current$set(label = "brier-by-forecaster")
tbl <- read_csv("../output/tables/brier-by-forecaster.csv")
knitr::kable(tbl, format = "latex", digits = 2, booktabs = TRUE,
             caption = "Average Brier score by forecaster group")
```

@@

Teasing out the chart and model effects is a bit harder through pairwise comparison. Instead, we directly coded whether a forecaster saw a chart or model forecast. Forcasters in conditions B and C and for questions that had a chart saw a chart, and forecasters in condition C looking at a question which had a machine forecast will have seen a model forecast. We also leave out the machine forecasts from the data, leaving almost 47,000 forecasts, of which about 25% were made with a chart available, and 21% also had a model forecast available.

We estimated two linear models to predict Brier scores for a forecast. The first had the specifiction:

\[
\textrm{Brier} \sim \alpha + \beta_1\textrm{Sees model} + \beta_2\textrm{Sees chart} + \beta_3\textrm{}
\]

The intercept $\alpha$ corresponds to the average Brier score for a Turker forecasting on a IFP that did not have chartable data, and who neither saw a chart nor machine forecast.

@@

\begin{figure}
\caption{\label{fig:model-chart-and-model-effects} Model estimates}
\includegraphics{../output/figures/model-chart-and-model-effects.png}
\end{figure}

\begin{table}
\caption{\label{tab:mdl-summary-stats}foo}
\centering
\begin{tabular}[t]{llr}
\toprule
Model & Statistic & Value\\
\midrule
Model 1: Linear model & N & 46956\\
 & R$^2$ & 0.013\\
 & Adj. R$^2$ & 0.013\\
Model 2: Linear model with IFP random intercepts & N & 46956\\
 & Marginal R$^2$ & 0.011\\
 & Conditional R$^2$ & 0.283\\
\bottomrule
\end{tabular}
\end{table}
